---
title: "Matrix Notebook – Schaum's Mastery"
author: "Rodolfo León"
format: 
  pdf:
    code-fold: true
    code-tools: true
---

# Matrix Notebook
This notebook contains matrices for exercises drawn from **Schaum's Matrix Operations** (and other Schaum volumes).

It serves as an **organized workbook** you can use for:
- LU decomposition (Crout, Doolittle, pivoting)
- Solving Ax = b
- Determinants
- Matrix inverses
- REF / RREF
- Rank
- Other linear algebra routines

All computations use the functions defined in:
functions/matrix_tools.R

```{r}
knitr::opts_knit$set(root.dir = "C:/Users/rodol/OneDrive/Documents/Schaums-mastery/matrices_R", echo = TRUE, warning = FALSE, message = TRUE)
```

```{r}
source("functions/matrix_tools.R")
library(pander)
panderOptions("digits", 6)
```

```{r}
# row_echelon_form()
# lu_c <- crout_decomposition()
# lu_d <- doolittle_decomposition()
# solve_lu()
# det_lu()
# matrix_inverse()
# rank_lu()
# matvec()
# crout_decomposition_pivoting()
# solv_doolittle()
# det_doolittle()
# inverse_doolittle()
# run_full_analysis() 
```


```{r}
# 3.3
A <- matrix(c(1, 2, -1,
              3, 8,  9,
              2, -1,  2),
            nrow = 3, byrow = TRUE)

lu_d <- doolittle_decomposition(A)
  pander("L (Doolittle)"); pander(lu_d$L)
  pander("U (Doolittle)"); pander(lu_d$U)
```

```{r}
# 3.4
G <- matrix(c(1,2,-2,3,
              -1, 1,0,2,
              3,-3,4,1,
              2,1,1,-2),
            nrow = 4, byrow = TRUE)

lu_d <- doolittle_decomposition(G)
  pander("L (Doolittle)"); pander(lu_d$L)
  pander("U (Doolittle)"); pander(lu_d$U)
```
```{r}
# 3.5
K <- matrix(c(1,2,-2,3,
              -1,-2,0,2,
              3,-3,0,1,
              2,1,1,-2),
            nrow = 4, byrow = TRUE)

lu_d <- doolittle_decomposition(K)
  pander("L (Doolittle)"); pander(lu_d$L)
  pander("U (Doolittle)"); pander(lu_d$U)
```

```{r}
F <- matrix(c(1,2,3,4,
              2,-2,-1,1,
              1,-3,4,-4,
              2,2,-3,4),
              nrow = 4, byrow = TRUE)

run_full_analysis(F)

```

```{r}
A <-  matrix(c(2,2,1,0,
               3,0,-1,1,
               0,1,0,5,
               -1,1,0,0), nrow=4, byrow=TRUE)

b <- c(10,-11,5,14)

solve_doolittle(A, b)

```

